Why does preregistration increase the persuasiveness of evidence? A Bayesian rationalization

### 2023-02-16 manuscript.Rmd
julias suggestions

---
title: "Why does preregistration increase the persuasiveness of evidence? A Bayesian rationalization."
bibliography: references.bib
abstract: "`r paste0(c('\\\\noindent', readLines('abstract.md')), collapse = ' ')`"
author:
  - name: Aaron Peikert
    affil: a, b, c
    email: peikert@mpib-berlin.mpg.de
  - name: Andreas M. Brandmaier
    affil: a, b, d 
    email: brandmaier@mpib-berlin.mpg.de
affiliation:
  - num: a
    address: |
      Center for Lifespan Psychology, Max Planck Institute for Human Development, Lentzeallee 94, 14195 Berlin, Germany
  - num: b
    address: |
     Max Planck UCL Centre for Computational Psychiatry and Ageing Research Berlin, Germany and London, UK
  - num: c
    address: |
      Department of Psychology, Humboldt-Universität zu Berlin, Unter den Linden 6, 10117 Berlin, Germany
  - num: d
    address: |
      MSB Medical School Berlin, Rüdesheimer Str. 50, 14197 Berlin, Germany
repro:
  scripts:
    - R/plots.R
  files:
    - abstract.md
    - data/simulation.rds
    - apa.csl
  packages:
    - here
    - tidyverse
    - ggplot2
    - ggthemes
    - patchwork
    - tikzDevice
    - rticles
    - gert
    - aaronpeikert/repro@5075336
  tlmgr:
    - pgf
    - preview
    - libertine
  apt:
    - git
    - rsync
    - curl
    - inkscape
output: rticles::tf_article
type: ARTICLE TEMPLATE
keywords: |
  preregistration; confirmation; exploration; hypothesis testing; bayesian; open science
header-includes: |
  \usepackage{hyperref}
  \usepackage[utf8]{inputenc}
  \def\tightlist{}
---

@@

<!--introduce confirmatory vs. exploratory-->
The scientific community has long pondered the vital distinction between exploration and confirmation, discovery and justification, hypothesis generation and hypothesis testing, or prediction and postdiction [@hoyningen-hueneContextDiscoveryContext2006; @shmueliExplainPredict2010; @nosekPreregistrationRevolution2018].
Despite the different names, it is fundamentally the same dichotomy that is at stake here.
There is a broad consensus that both approaches are necessary for science to progress; exploration for making new discoveries and confirmation for exposing these discoveries to potential falsification and assess in how far they have been corroborated.
However, mistaking exploratory findings for empirically confirmed results is dangerous.
It inflates the likelihood of believing that there is evidence supporting a given explanation even if it is false.
A variety of problems, such as researchers' degrees of freedom together with researchers' hindsight bias or naive p-hacking have led to such mistakes becoming commonplace yet unnoticed for a long time.
Recognizing them has led to a crisis of confidence in the empirical sciences [@ioannidisWhyMostPublished2005], and psychology in particular [@opensciencecollaborationEstimatingReproducibilityPsychological2015].
As a response to the crisis, more and more researchers preregister their hypotheses and their data collection and analysis plans in advance of the study [@nosekPreregistrationRevolution2018].
They do so to stress the predictive nature of their registered statistical analyses; often with the hopes of obtaining a label that marks the study as "confirmatory".
Indeed, rigorous application of preregistration prevents researchers from reporting a set of results produced by an arduous process of trial and error as a simple confirmatory story [@wagenmakersAgendaPurelyConfirmatory2012] while keeping low false-positive rates.
This promise of a clear distinction between confirmation and exploration has obvious appeal to many who already accepted the practice.
Still, the majority of empirical researchers do not routinely preregister their studies.
One reason may be that some do not find that the theoretical advantages outweigh the practical hurdles, such as specifying every aspect of a theory in advance.
We believe that we can reach a greater acceptance of preregistration by explicating a more general objective of preregistration that benefits all kinds of studies, even those that allow data-dependent decisions.

<!--the current goal of preregistration is confirmation-->
One goal of preregistration that has received widespread attention, is to clearly distinguish confirmatory from exploratory research [@mellorEasyPreregistrationWill2018; @nosekPreregistrationRevolution2018; @wagenmakersAgendaPurelyConfirmatory2012; @simmonsPreregistrationWhyHow2021; @bakkerEnsuringQualitySpecificity2020].
In such a narrative, preregistration is justified by a confirmatory research agenda.
However, two problems become apparent under closer inspection.
First, many researcher do not subscribe to a purely confirmatory research agenda.
Second, it is not necessary to strictly map the categories preregistered vs. non-preregistered onto the categories confirmatory vs. exploratory research.

@@

<!---many researcher do not subscribe to confirmatory agenda--->
We argue that such criticism is not directed against preregistration itself but against a justification through a confirmatory research agenda [@wagenmakersAgendaPurelyConfirmatory2012].
When researchers criticize preregistration as being too inflexible to fit their research question, they often simply acknowledge that their research goals are not strictly confirmatory.
Forcing researchers into adopting a strictly confirmatory research agenda does not only imply changing *how* they investigate a phenomenon but also *what* research questions they ask.
However reasonable such a move is, changing core believes of a large community is much harder than convincing them that a method is well justified.
We therefore attempt to disentangle the *methodological* goals of preregistration from the *ideological* goals of confirmatory science.
It is likely that psychology needs more confirmatory studies to progress as a science, but quite independently of this, preregistration can be useful for any kind of study on the continuum between strictly confirmatory and fully exploratory.

<!---outline-structure--->
To form such an objective for preregistration, we first introduce some tools of Bayesian philosophy of science and map the exploration/confirmation distinction onto a dimensional quantity we call "theoretical risk" [a term borrowed from @meehlTheoreticalRisksTabular1978, but formalized as the probability of proving a hypothesis wrong, if it does not hold], which is inversely related to the type-I error rate in null hypothesis testing.

Further, we outline two interpretations of how preregistration impacts theoretical risk.
The first interpretation corresponds to the traditional application of preregistration to research paradigms that focus on confirmation by maximizing the theoretical risk or equivalently by limiting type-I error (when dichotomous decisions about theories are an inferential goal).
The second interpretation is our main contribution and demonstrates the broad applicability of preregistration to both exploratory and confirmatory studies that are implemented as preregistered or have undergone changes after preregistration.
We argue that the classic view on the utility of preregistration can be interpreted as maximization of theoretical risk, which is reduced by researchers' degrees of freedom, p-hacking, and such.
Importantly, we argue that theoretical risk is not necessarily directly maximized by preregistration, but rather the uncertainty in judging the theoretical risk is minimized.

@@

The assumption of rationality can be connected to Bayesian reasoning and leads to our adoption of the framework.
Our rationale is as follows.
Researchers who decide to conduct a certain study are actually choosing a study to bet on.
They have to "place the bet" by conducting the study, therefore, invest resources and stand to gain epistemic value with some probability.
This conceptualization of choosing a study as a betting problem allows us to apply a "Dutch Book" argument [@christensenCleverBookiesCoherent1991].
This argument states that any better must follow the axioms of probability to avoid being "irrational," i.e., accepting bets that lead to sure losses.
Fully developing a Dutch book argument for this problem requires careful consideration of what kind of studies to include as possible bets, defining a conversion rate from the stakes to the reward, and modelling what liberties researchers have in what studies to conduct.
Without deliberating these concepts further, we find it persuasive that researchers should not violate the axioms of probability if they have some expectation about what they stand to gain with some likelihood from conducting a study.
The axioms of probability are sufficient to derive the Bayes formula, on which we will heavily rely for our further arguments.
The argument is not sufficient, however, to warrant conceptualizing the kind of epistemic value we reason about in terms of posterior probability; that remains a leap of faith.
However, the argument applies to any reward function that satisfies the "statistical relevancy condition" [@wesleyc.salmonStatisticalExplanation1970; @fetzerStatisticalExplanations1974].
That is, evidence only increases epistemic value for a theory, if the evidence is more likely to observe under the theory than without it.

@@

Let us make some simplifying assumptions and define notation.
We restrict ourselves to evidence of a binary nature (either it was observed or not) since continuous evidence would lead to some quite involved derivations.
We denote the probability of a hypothesis before observing evidence as $P(H)$ and its complement as $P(\neg{}H) = 1 - P(H)$.
The probability of observing evidence under some hypothesis is $P(E|H)$.
We can calculate the probability of the hypothesis after observing the evidence with the help of the Bayes formula:

@@

In short, we want to increase posterior probability $P(H|E)$.
Increases in posterior probability $P(H|E)$ are associated with increased epistemic value, for which we want to maximize the expectation.
So how can we increase posterior probability?
The Bayes formula yields three components that influence corroboration, namely $P(H)$, $P(E|H)$ and $P(E)$.
The first option leads us to the unsurprising conclusion that higher a priori probability $P(H)$ leads to higher posterior probability $P(H|E)$.
If a hypothesis is more probable to begin with, observing evidence in favor will result a more corroborated hypothesis, all else equal.
However, the prior probability of a hypothesis is nothing our study design can change.
The second option is similarly commonsensical; that is, an increase in $P(E|H)$ leads to higher posterior probability $P(H|E)$.
$P(E|H)$ is the probability of obtaining evidence for a theory, when the theory holds.
We call this probability of detecting evidence given that the theory holds "detectability."
Consequently, researchers should ensure that their study design allows them to find evidence for their hypothesis, in case it is true.
When applied strictly within the bounds of null hypothesis testing, detectability is equivalent to power (or the inverse of type-II error rate).
However, while detectability is of great importance for study design, it is not directly relevant to the objective of preregistration.
Thus, $P(E)$ remains to be considered.
Since $P(E)$ is the denominator, increasing it will decrease the posterior probability: The more unlikely it is to observe evidence, the more it increases the probability of the hypothesis if we do observe it.
In other words, high risk, high reward.

If we equate riskiness with a low probability of obtaining evidence, the Bayesian rationale perfectly aligns with the observation that risky predictions lead to persuasive evidence.
This tension between high risk leading to high reward is central to our consideration of preregistration.
A high-risk, high-reward strategy is bound to result in many losses that are eventually absorbed by the high gains.
Sustaining many "failed" studies is not exactly aligned with the incentive structure under which many, if not most, researchers operate.
Consequently, researchers have an incentive to appear to take more risks than they actually do, which misleads their readers to give their claims more credence than they deserve.
It is at this juncture that the practice and mispractice of preregistration comes into play.
We argue that the main function of preregistration is to enable proper judgment of the riskiness of a study.

To better understand how preregistrations can achieve that, let us take a closer look at what factors contribute to $P(E)$. 
Using the law of total probability, we can split $P(E)$ into two terms:

@@

We already have noted that there is not much to do about prior probability ($P(H)$, and hence its counter probability $P(\neg{}H)$), and that it is common sense to increase detectability $P(E|H)$.
The real lever to pull is, therefore, $P(E|\neg{}H)$.
This probability tells us how likely it is that we find evidence in favor of the theory when in fact, the theory is not true.
Its counter probability $P(\neg{}E|\neg{}H)= 1 - P(E|\neg{}H)$ is what we call "theoretical risk", because it is the risk a theory takes on in predicting the occurrence of particular evidence in its favour.
We "borrow" the term from @meehlTheoreticalRisksTabular1978, though he has not assigned it to the probability $P(\neg{}E|\neg{}H)$.
@kuklaClinicalStatisticalTheory1990 argued that the core arguments in @meehlAppraisingAmendingTheories1990 can be reconstructed in purely Bayesian framework.
However, while he did not name $P(\neg{}E|\neg{}H)$ but did suggest that @meehlTheoreticalRisksTabular1978 used the term "very strange coincidence" for a small $P(E|\neg{}H)$ which would imply that $P(\neg{}E|\neg{}H)$ can be related to or even equated to theoretical risk.

@@

The last statement sounds like a truism but is directly related to Popper's seminal criterion of demarcation.
He stated that if it is impossible to prove that a hypothesis is false ($P(\neg{}E|\neg{}H) = 0$, theoretical risk is zero), it cannot be considered a scientific hypothesis [@popperLogicScientificDiscovery2002, p. 18].
We note these relations to underline that the Bayesian rational we apply here is able to reconstruct many commonly held views on riskiness and epistemic value.

Both theoretical risk $P(\neg{}E|\neg{}H)$ and detectability $P(E|H)$ aggregate uncountable influences, otherwise they could not model the process of evidential support for theories.
To illustrate the concepts we introduced up to here, consider the following  example of a single theory and three experiments that may test it. 
The experiments were created to illustrate how they may differ in their theoretical risk and detectability.
Suppose the primary theory is about the cognitive phenomenon of "insight."
For the purpose of illustration, we define it, with quite some hand-waving, as a cognitive abstraction that allows agents to consistently solve a well-defined class of problems.
We present the hypothesis that the following problem belongs to such a class of insight problems:

@@

1. The experimenter gives a hint that  the problem is easy to solve when using Roman numerals; if all students come up with the solution, she records it as evidence for the hypothesis.
2. The experimenter shows the solution "IX" and explains it; if all students come up with the solution, she records it as evidence for the hypothesis.
3. The experimenter does nothing; if all students come up with the solution, she records it as evidence for the hypothesis.

We argue that experiment 1 has high theoretical risk $P(\neg{}E_1|\neg{}H)$ and high detectability $P(E_1|H)$.
If "insight" has nothing to do with solving the problem ($\neg{}H$), then presenting the insight that Roman numerals could be used, should not lead to all students solving the problem ($\neg{}E_1$); the experiment therefore has high theoretical risk $P(\neg{}E_1|\neg{}H)$.
Conversely, if insight is required to solve the problem ($H$), then it is likely to help all students to solve the problem ($E_1$), the experiment therefore has high detectability $P(E_1|H)$.
The second experiment, on the other hand, has low theoretical risk $P(\neg{}E_2|\neg{}H)$.
Even if "insight" has nothing to do with solving the problem ($\neg{}H$), there are other plausible reasons for observing the evidence ($E_2$), because the students could simply copy the solution, without having any insight.
With regard to detectability, experiments 1 and 2 differ in no obvious way.
Experiment 3, however, also has low detectability.
It is unlikely that all students come up with the correct solution in a short time ($E_3$), even if insight is required ($H$) experiment 3 therefore has low detectability $P(E_3|H)$.
The theoretical risk, however, is also low in absolute terms, but high compared to the detectability (statistical relevancy condition is satisfied).
In the unlikely event that all 10 students place their matches to form the Roman numeral VIII ($E_3$), it is probably due to insight ($H$) and not by chance $P(\neg{}E_3|\neg{}H)$).
Of course, in practice, we would allow the evidence to be probabilistic, e.g., relax the requirement of "all students" to nine out of ten students, more than eight, and so forth.

As argued earlier, the remainder of the paper will focus on binary, non-probabilisitic evidence to keep the mathematical notation as simple as possible.
We discuss the relation between statistical methods and theoretical risk in the [Statistical Methods] section.

@@

First, the theory itself limits theoretical risk:
Some theories simply do not make risky predictions, and preregistration will not change that.
Consider the case of a researcher contemplating the relation between two sets of variables.
Suppose each set is separately well studied, and strong theories tell the researcher how the variables within the set relate.
However, our imaginary researcher now considers the relation between these two sets.
For lack of a better theory, they assume that some relation between any variables of the two sets exists.
This is not a risky prediction to make in psychology, even without statistical issues like alpha inflation [@orbenCrudReDefined2020].
However, we would consider it a success if the researcher would use the evidence to develop a more precise (and therefore risky) theory, e.g., by specifying which variables from one set relate to which variables from the other set, to what extent, in which direction, with which functional shape, etc.
We will later show that preregistration increases the degree of belief in the further specified theory, though it remains low till being substantiated by testing it again.
The point, however, is that we want to show that preregistration increases the expected epistemic value regardless of the theory being tested.

@@

Statistical methods stand out among these factors because we have a large and well-understood toolbox for assessing and controlling their contribution to theoretical risk.
Examples of our ability to exert this control are the choice of type-I error rate, adjustments for multiple testing, the use of corrected fit measures (i.e., adjusted R²), information criteria, or cross-validation in machine learning.
These tools help us account for biases in statistical methods that increase the likelihood of signifying results even when there are none $P(E|\neg{}H)$.

@@

As we noted, it is possible to quantify how statistical models affect the theoretical risk based on mathematical considerations and simulation.
However, other factors in the broader context of the study are much harder to quantify.
If one chooses to focus only on the contribution of statistical methods to theoretical risk, one is bound to overestimate it.
Take, for example, a t-test  of mean differences in two samples.
Under ideal circumstances (assumption of independence, normality of residuals, equal variance), it stays true to its type-I error rate.
However, researchers might do many very reasonable things in the broader context of the study that affect theoretical risk:
They might exclude outliers, choose to drop an item before computing a sum score, broaden, enlarge their definition of the population to be sampled, translate their questionnaires into a different language, impute missing values, switch between different estimators of the pooled variance, or any number of other things.
All of these decisions carry a small risk that they increase the likelihood of obtaining evidence despite the underlying research hypothesis being false.
Even if the t-test itself perfectly maintains its type I error rate, these factors influence $P(E|\neg{}H)$.
While, in theory, these factors may leave $P(E|\neg{}H)$ unaffected or even decrease it, we argue that this is not the case in practice.
Whether researchers want to or not, they continuously process information about how the study is going, except under strict blinding.
While one can hope that processing this information does not affect their decision making either way, this cannot be secured.
We, therefore, conclude that statistical properties only guarantee a lower bound for theoretical risk.
The only thing we can conclude with some certainty is that theoretical risk is not higher than what the statistical model guarantees without knowledge about the other factors at play.

@@

Where $\tau{}$ is the upper bound of theoretical risk (e.g., 0.95 for a statistical model with a nominal type-I error rate of 5%).
We chose this uniform distribution to capture our statement that a statistical model only guarantees an upper bound to theoretical risk (and since it is a probability the lower bound is 0) as it is the maximum entropy distribution under this assumption and conforms therefore to our Bayesian framework [@giffinUpdatingProbabilitiesData2007].

@@

```{r posterior-fig, dev = "tikz", cache=TRUE, warning=FALSE}
#| fig.cap: "\\label{fig:posterior}Posterior Probability (corroboration as firmness) as a function of theoretical risk $\\tau$, where $\\tau$ is either certain (solid line) or maximally uncertain (dotted line)."

>>>>

---
title: "Why does preregistration increase the persuasiveness of evidence? A Bayesian rationalization"
bibliography: references.bib
abstract: "`r paste0(c('\\\\noindent', readLines('abstract.md')), collapse = ' ')`"
author:
  - name: Aaron Peikert
    affil: a, b, c
    email: peikert@mpib-berlin.mpg.de
  - name: Andreas M. Brandmaier
    affil: a, b, d 
    email: brandmaier@mpib-berlin.mpg.de
affiliation:
  - num: a
    address: |
      Center for Lifespan Psychology, Max Planck Institute for Human Development, Lentzeallee 94, 14195 Berlin, Germany
  - num: b
    address: |
     Max Planck UCL Centre for Computational Psychiatry and Ageing Research Berlin, Germany and London, UK
  - num: c
    address: |
      Department of Psychology, Humboldt-Universität zu Berlin, Unter den Linden 6, 10117 Berlin, Germany
  - num: d
    address: |
      MSB Medical School Berlin, Rüdesheimer Str. 50, 14197 Berlin, Germany
repro:
  scripts:
    - R/plots.R
  files:
    - abstract.md
    - data/simulation.rds
    - apa.csl
  packages:
    - here
    - tidyverse
    - ggplot2
    - ggthemes
    - patchwork
    - tikzDevice
    - rticles
    - gert
    - aaronpeikert/repro@5075336
  tlmgr:
    - pgf
    - preview
    - libertine
  apt:
    - git
    - rsync
    - curl
    - inkscape
output: rticles::tf_article
type: ARTICLE TEMPLATE
keywords: |
  preregistration; confirmation; exploration; hypothesis testing; Bayesian; open science
header-includes: |
  \usepackage{hyperref}
  \usepackage[utf8]{inputenc}
  \def\tightlist{}
---

@@

<!--introduce confirmatory vs. exploratory-->
The scientific community has long pondered the vital distinction between exploration and confirmation, discovery and justification, hypothesis generation and hypothesis testing, or prediction and postdiction  [@hoyningen-hueneContextDiscoveryContext2006; @shmueliExplainPredict2010; @nosekPreregistrationRevolution2018].
Despite the different names, it is fundamentally the same dichotomy that is at stake here.
There is a broad consensus that both approaches are necessary for science to progress; exploration, to make new discoveries and confirmation, to expose these discoveries to potential falsification and assess in how far the theory has been corroborated.
However, mistaking exploratory findings for empirically confirmed results is dangerous.
It inflates the likelihood of believing that there is evidence supporting a given explanation even if it is false.
A variety of problems, such as researchers' degrees of freedom together with researchers' hindsight bias or naive p-hacking have led to such mistakes becoming commonplace yet unnoticed for a long time.
Recognizing them has led to a crisis of confidence in the empirical sciences [@ioannidisWhyMostPublished2005], and psychology in particular [@opensciencecollaborationEstimatingReproducibilityPsychological2015].
As a response to the crisis, evermore researchers preregister their hypotheses and their data collection and analysis plans in advance of their studies [@nosekPreregistrationRevolution2018].
They do so to stress the predictive nature of their registered statistical analyses; often with the hopes of obtaining a label that marks the study as "confirmatory".
Indeed, rigorous application of preregistration prevents researchers from reporting a set of results produced by an arduous process of trial and error as a simple confirmatory story [@wagenmakersAgendaPurelyConfirmatory2012] while keeping low false-positive rates.
This promise of a clear distinction between confirmation and exploration has obvious appeal to many who have already accepted the practice.
Still, the majority of empirical researchers do not routinely preregister their studies.
One reason may be that some do not find that the theoretical advantages outweigh the practical hurdles, such as specifying every aspect of a theory in advance.
We believe that we can reach a greater acceptance of preregistration by explicating a more general objective of preregistration that benefits all kinds of studies, even those that allow data-dependent decisions.

<!--the current goal of preregistration is confirmation-->
One goal of preregistration that has received widespread attention is to clearly distinguish confirmatory from exploratory research [@mellorEasyPreregistrationWill2018; @nosekPreregistrationRevolution2018; @wagenmakersAgendaPurelyConfirmatory2012; @simmonsPreregistrationWhyHow2021; @bakkerEnsuringQualitySpecificity2020].
In such a narrative, preregistration is justified by a confirmatory research agenda.
However, two problems become apparent under closer inspection.
First, many researchers do not subscribe to a purely confirmatory research agenda.
Second, it is not necessary to strictly map the categories preregistered vs. non-preregistered onto the categories confirmatory vs. exploratory research.

@@

<!---many researcher do not subscribe to confirmatory agenda--->
We argue that such criticism is not directed against preregistration itself but against a justification through a confirmatory research agenda [@wagenmakersAgendaPurelyConfirmatory2012].
When researchers criticize preregistration as being too inflexible to fit their research question, they often simply acknowledge that their research goals are not strictly confirmatory.
Forcing researchers into adopting a strictly confirmatory research agenda does not only imply changing *how* they investigate a phenomenon but also *what* research questions they pose.
However reasonable such a move is, changing core beliefs of a large community is much harder than convincing them that a method is well justified.
We therefore attempt to disentangle the *methodological* goals of preregistration from the *ideological* goals of confirmatory science.
It is likely that psychology needs more confirmatory studies to progress as a science, but quite independently of this, preregistration can be useful for any kind of study on the continuum between strictly confirmatory and fully exploratory.

<!---outline-structure--->
To form such an objective for preregistration, we first introduce some tools of Bayesian philosophy of science and map the exploration/confirmation distinction onto a dimensional quantity we call "theoretical risk" [a term borrowed from @meehlTheoreticalRisksTabular1978, but formalized as the probability of proving a hypothesis wrong if it does not hold], which is inversely related to the type-I error rate in null hypothesis testing.

Further, we outline two interpretations of how preregistration impacts theoretical risk.
The first interpretation corresponds to the traditional application of preregistration to research paradigms that focus on confirmation by maximizing the theoretical risk or equivalently by limiting type-I error (when dichotomous decisions about theories are an inferential goal).
The second interpretation is our main contribution and demonstrates the broad applicability of preregistration to both exploratory and confirmatory studies that are implemented as preregistered or have undergone changes after preregistration.
We argue that the classic view on the utility of preregistration can be interpreted as maximization of theoretical risk, which is reduced by researchers' degrees of freedom, p-hacking, and suchlike.
Importantly, we argue that theoretical risk is not necessarily directly maximized by preregistration, but rather that the uncertainty in judging the theoretical risk is minimized.

@@

The assumption of rationality can be connected to Bayesian reasoning and leads to our adoption of the framework.
Our rationale is as follows.
Researchers who decide to conduct a certain study are actually choosing a study to bet on.
They have to "place the bet" by conducting the study, therefore, invest resources and stand to gain epistemic value with some probability.
This conceptualization of choosing a study as a betting problem allows us to apply a "Dutch book" argument [@christensenCleverBookiesCoherent1991].
This argument states that any better must follow the axioms of probability to avoid being "irrational," i.e., accepting bets that lead to sure losses.
Fully developing a Dutch book argument for this problem requires careful consideration of what kind of studies to include as possible bets, defining a conversion rate from the stakes to the reward, and modelling what liberties researchers have in what studies to conduct.
Without deliberating these concepts further, we find it persuasive that researchers should not violate the axioms of probability if they have some expectation about what they stand to gain with some likelihood from conducting a study.
The axioms of probability are sufficient to derive the Bayes formula, on which we will heavily rely for our further arguments.
The argument is not sufficient, however, to warrant conceptualizing the kind of epistemic value we reason about in terms of posterior probability; that remains a leap of faith.
However, the argument applies to any reward function that satisfies the "statistical relevancy condition" [@wesleyc.salmonStatisticalExplanation1970; @fetzerStatisticalExplanations1974].
That is, evidence only increases epistemic value for a theory, if the evidence is more likely to observe under the theory than without it.

@@

Let us make some simplifying assumptions and define our notation.
We restrict ourselves to evidence of a binary nature (either it was observed or not) since continuous evidence would lead to some quite involved derivations.
We denote the probability of a hypothesis before observing evidence as $P(H)$ and its complement as $P(\neg{}H) = 1 - P(H)$.
The probability of observing evidence under some hypothesis is $P(E|H)$.
We can calculate the probability of the hypothesis after observing the evidence with the help of the Bayes formula:

@@

In short, we want to increase posterior probability $P(H|E)$.
Increases in posterior probability $P(H|E)$ are associated with increased epistemic value, for which we want to maximize the expectation.
So how can we increase posterior probability?
The Bayes formula yields three components that influence corroboration, namely $P(H)$, $P(E|H)$ and $P(E)$.
The first option leads us to the unsurprising conclusion that higher a priori probability $P(H)$ leads to higher posterior probability $P(H|E)$.
If a hypothesis is more probable to begin with, observing evidence in its favor will result a more corroborated hypothesis, all else being equal.
However, the prior probability of a hypothesis is nothing our study design can change.
The second option is similarly commonsensical; that is, an increase in $P(E|H)$ leads to higher posterior probability $P(H|E)$.
$P(E|H)$ is the probability of obtaining evidence for a theory, when the theory holds.
We call this probability of detecting evidence given that the theory holds "detectability."
Consequently, researchers should ensure that their study design allows them to find evidence for their hypothesis, in case it is true.
When applied strictly within the bounds of null hypothesis testing, detectability is equivalent to power (or the inverse of type-II error rate).
However, while detectability is of great importance for study design, it is not directly relevant to the objective of preregistration.
Thus, $P(E)$ remains to be considered.
Since $P(E)$ is the denominator, increasing it will decrease the posterior probability: The more unlikely it is to observe evidence, the more it increases the probability of the hypothesis if we do observe it.
In other words, high risk, high gain.

If we equate riskiness with a low probability of obtaining evidence, the Bayesian rationale perfectly aligns with the observation that risky predictions lead to persuasive evidence.
This tension between high risk leading to high gain is central to our consideration of preregistration.
A high-risk, high-gain strategy is bound to result in many losses that are eventually absorbed by the high gains.
Sustaining many "failed" studies is not exactly aligned with the incentive structure under which many, if not most, researchers operate.
Consequently, researchers have an incentive to appear to take more risks than they actually do, which misleads their readers to give their claims more credence than they deserve.
It is at this juncture that the practice and mispractice of preregistration comes into play.
We argue that the main function of preregistration is to enable proper judgment of the riskiness of a study.

To better understand how preregistrations can achieve that, let us take a closer look at the factors contributing to $P(E)$. 
Using the law of total probability, we can split $P(E)$ into two terms:

@@

We have already noted that there is not much to be done about prior probability ($P(H)$, and hence its counter probability $P(\neg{}H)$), and that it is common sense to increase detectability $P(E|H)$.
The real lever to pull is therefore $P(E|\neg{}H)$.
This probability tells us how likely it is that we find evidence in favor of the theory when in fact, the theory is not true.
Its counter probability $P(\neg{}E|\neg{}H)= 1 - P(E|\neg{}H)$ is what we call "theoretical risk", because it is the risk a theory takes on in predicting the occurrence of particular evidence in its favour.
We "borrow" the term from @meehlTheoreticalRisksTabular1978, though he has not assigned it to the probability $P(\neg{}E|\neg{}H)$.
@kuklaClinicalStatisticalTheory1990 argued that the core arguments in @meehlAppraisingAmendingTheories1990 can be reconstructed in a purely Bayesian framework.
However, while he did not mention $P(\neg{}E|\neg{}H)$ he suggested that @meehlTheoreticalRisksTabular1978 used the term "very strange coincidence" for a small $P(E|\neg{}H)$ which would imply, that $P(\neg{}E|\neg{}H)$ can be related to or even equated to theoretical risk.

@@

The last statement sounds like a truism but is directly related to Popper's seminal criterion of demarcation.
He stated that if it is impossible to prove that a hypothesis is false ($P(\neg{}E|\neg{}H) = 0$, theoretical risk is zero), it cannot be considered a scientific hypothesis [@popperLogicScientificDiscovery2002, p. 18].
We note these relations to underline that the Bayesian rationale we apply here is able to reconstruct many commonly held views on riskiness and epistemic value.

Both theoretical risk $P(\neg{}E|\neg{}H)$ and detectability $P(E|H)$ aggregate countless influences, otherwise they could not model the process of evidential support for theories.
To illustrate the concepts we have introduced up to here, consider the following  example of a single theory and three experiments that may test it. 
The experiments were created to illustrate how they may differ in their theoretical risk and detectability.
Suppose the primary theory is about the cognitive phenomenon of "insight."
For the purpose of illustration, we define it, with quite some hand-waving, as a cognitive abstraction that allows agents to consistently solve a well-defined class of problems.
We present the hypothesis that the following problem belongs to such a class of insight problems:

@@

1. The experimenter gives a hint that  the problem is easy to solve when using Roman numerals; if all students come up with the solution, she records it as evidence for the hypothesis.
2. The experimenter shows the solution "VIII" and explains it; if all students come up with the solution, she records it as evidence for the hypothesis.
3. The experimenter does nothing; if all students come up with the solution, she records it as evidence for the hypothesis.

We argue that experiment 1 has high theoretical risk $P(\neg{}E_1|\neg{}H)$ and high detectability $P(E_1|H)$.
If "insight" has nothing to do with solving the problem ($\neg{}H$), then presenting the insight that Roman numerals can be used, should not lead to all students solving the problem ($\neg{}E_1$); the experiment therefore has high theoretical risk $P(\neg{}E_1|\neg{}H)$.
Conversely, if insight is required to solve the problem ($H$), then it is likely to help all students to solve the problem ($E_1$), the experiment therefore has high detectability $P(E_1|H)$.
The second experiment, on the other hand, has low theoretical risk $P(\neg{}E_2|\neg{}H)$.
Even if "insight" has nothing to do with solving the problem ($\neg{}H$), there are other plausible reasons for observing the evidence ($E_2$), because the students could simply copy the solution, without having any insight.
With regard to detectability, experiments 1 and 2 differ in no obvious way.
Experiment 3, however, also has low detectability.
It is unlikely that all students will come up with the correct solution in a short time ($E_3$), even if insight is required ($H$); experiment 3 therefore has low detectability $P(E_3|H)$.
The theoretical risk, however, is also low in absolute terms, but high compared to the detectability (statistical relevancy condition is satisfied).
In the unlikely event that all 10 students place their matches to form the Roman numeral VIII ($E_3$), it is probably due to insight ($H$) and not by chance $P(\neg{}E_3|\neg{}H)$).
Of course, in practice, we would allow the evidence to be probabilistic, e.g., relax the requirement of "all students" to nine out of ten students, more than eight, and so forth.

As mentioned earlier, the remainder of the paper will focus on binary, non-probabilisitic evidence to keep the mathematical notation as simple as possible.
We discuss the relation between statistical methods and theoretical risk in the [Statistical Methods] section.

@@

First, the theory itself limits theoretical risk:
Some theories simply do not make risky predictions, and preregistration will not change that.
Consider the case of a researcher contemplating the relation between two sets of variables.
Suppose each set is separately well studied, and strong theories tell the researcher how the variables within the set relate.
However, our imaginary researcher now considers the relation between these two sets.
For lack of a better theory, they assume that some relation between any variables of the two sets exists.
This is not a risky prediction to make in psychology, even without statistical issues like alpha inflation [@orbenCrudReDefined2020].
However, we would consider it a success if the researcher would use the evidence to develop a more precise (and therefore risky) theory, e.g., by specifying which variables from one set relate to which variables from the other set, to what extent, in which direction, with which functional shape, etc.
We will later show that preregistration increases the degree of belief in the further specified theory, though it remains low till being substantiated by testing the theory again.
The point, however, is that we want to show that preregistration increases the expected epistemic value regardless of the theory being tested.

@@

Statistical methods stand out among these factors because we have a large and well-understood toolbox for assessing and controlling their contribution to theoretical risk.
Examples of our ability to exert this control are the choice of type-I error rate, adjustments for multiple testing, the use of corrected fit measures (i.e., adjusted R²), information criteria, or cross-validation in machine learning.
These tools help us account for biases in statistical methods that decrease theoretical risk (and its inverse, $P(E|\neg{}H$).

@@

As we have noted, it is possible to quantify how statistical models affect the theoretical risk based on mathematical considerations and simulation.
However, other factors in the broader context of a study are much harder to quantify.
If one chooses to focus only on the contribution of statistical methods to theoretical risk, one is bound to overestimate it.
Take, for example, a t-test  of mean differences in two samples.
Under ideal circumstances (assumption of independence, normality of residuals, equal variance), it stays true to its type-I error rate.
However, researchers may do many very reasonable things in the broader context of the study that affect theoretical risk:
They might exclude outliers, choose to drop an item before computing a sum score, broaden, enlarge their definition of the population to be sampled, translate their questionnaires into a different language, impute missing values, switch between different estimators of the pooled variance, or any number of other things.
All of these decisions carry a small risk that they will increase the likelihood of obtaining evidence despite the underlying research hypothesis being false.
Even if the t-test itself perfectly maintains its type I error rate, these factors influence $P(E|\neg{}H)$.
While, in theory, these factors may leave $P(E|\neg{}H)$ unaffected or even decrease it, we argue that this is not the case in practice.
Whether researchers want to or not, they continuously process information about how the study is going, except under strict blinding.
While one can hope that processing this information does not affect their decision making either way, this cannot be ascertained.
Therefore, we conclude that statistical properties only guarantee a lower bound for theoretical risk.
The only thing we can conclude with some certainty is that theoretical risk is not higher than what the statistical model guarantees without knowledge about the other factors at play.

@@

Where $\tau{}$ is the upper bound of theoretical risk (e.g., 0.95 for a statistical model with a nominal type-I error rate of 5%).
We chose this uniform distribution to capture our statement that a statistical model only guarantees an upper bound to theoretical risk (and since it is a probability the lower bound is 0) as it is the maximum entropy distribution under this assumption and hence conforms to our Bayesian framework [@giffinUpdatingProbabilitiesData2007].

@@

```{r posterior-fig, dev = "tikz", cache=TRUE, warning=FALSE}
#| fig.cap: "\\label{fig:posterior}Posterior probability (corroboration as firmness) as a function of theoretical risk $\\tau$, where $\\tau$ is either certain (solid line) or maximally uncertain (dotted line)."
